---
title: "Biol 3872 Lab5-Assignment"
output: html_document
date: "`r Sys.Date()`"
author: ""
student_id: "lord jessington dunnford"
---

**BEFORE STARTING:**\
1. Include your name (as shown in Brightspace) in the `author` section above as text (e.g., "Liam MacNeil") and Banner ID ("B00123456") in `student_id`.\
2. Install and load the packages required to run your code chunks/

Assignments will largely be evaluated based on your code chunk answers. If they are incomplete or formatted incorrectly (e.g., code chunks don't run), no marks will be awarded. I recommend experimenting with solving these answers in a separate script/markdown document and reporting your answers with minimal script.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}
## Task 1

A)  Define an object `Euler` containing Euler's number (`exp(1)`). Write an `if` statement which tests if this value is positive and if so, create a new object called `growth` which calculates growth amount of a fish weighing 10 kg (P), at a 5% (0.05) growth rate each year (r), over two years (t).

**Growth equation** = P \* e^r \* t^

```{r}
# YOUR CODE HERE
Euler <- exp(1)

if (Euler >= 0){
  size <- 10 * exp(0.05*2)
  growth <- size - 10
}

growth
```
:::

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}
## Task 2

A)  Define an object `option1` equal to 9, and `option2` equal to 13. Write an explicit `ifelse` statement (not a vectorized version) which tests if one is bigger than the other and if `TRUE`, store this object name in a new variable called `which_max`.

```{r}
# YOUR CODE HERE
option1 <- 9
option2 <- 13

ifelse(option1 > option2, whichmax <- "option1", whichmax <- "option2")
```
:::

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}
## Task 3

A)  Using the `marine_survey` data frame, create a column called `environments` using an `ifelse` statement with two conditions: (1) `if` `species_count` is greater than the upper third quantile of `species_count` **and** (*&*) `if` `temperature_c` is greater than the mean temperature, call this row "diverse & hot"- if it's not `TRUE`, call this row "not"

**HINT** You can use a vectorized `ifelse` to achieve this. To grab the upper third quantile, check `?quantile` and the `probs =` argument.

B)  Below I calculate the median value of each numeric column where `environments` are "diverse & hot" and store it in a vector called `diverse_hot_envs`. Fix this code to run properly.

```{r}
# Create a data frame
marine_survey <- data.frame(
  station = c("ST1", "ST2", "ST3", "ST4", "ST5", "ST6"),
  depth_m = c(5, 15, 25, 10, 30, 20),
  temperature_c = c(22.5, 18.3, 12.8, 20.1, 10.5, 15.2),
  salinity_ppt = c(35.2, 35.8, 36.1, 35.5, 36.5, 35.9),
  species_count = c(45, 32, 18, 38, 12, 25)
)

# A) 

# YOUR CODE HERE
#find cutoff value for upper third of species_count
upper_third <- quantile(marine_survey$species_count, probs = 2/3)
#find mean temperature
mean_temp <- mean(marine_survey$temperature_c)

#create environments column, with contents depending on whether species_count is
#in the upper third quantile
marine_survey$environments <- ifelse(marine_survey$species_count > upper_third &
                                     marine_survey$temperature_c > mean_temp, 
                                     'diverse & hot', 'not')

# B)
# CODE TO FIX

diverse_hot_envs <- sapply(marine_survey[, 2:5], function(x) {
median(ifelse(marine_survey$environments == "diverse & hot", x, NA),
      na.rm = TRUE)
})

#original for reference:
#numeric_columns <- sapply(marine_survey, as.numeric)

# diverse_hot_envs <- sapply(marine_survey[, numeric_columns], function(x) {
# median(ifelse(marine_survey$environments == "diverse & hot", q, NA),
#       na.rm = FALSE)
# })


```
:::

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}
## Task 4

A)  Generate a vector called `sheep` of length 100 counting by 1, from 1 to 100.

B)  Write a `for()` loop which multiples all elements of `sheep` by pi and store this in a vector called `counting_sheep`. *Remember you that need an empty vector to fill and that "[]" lets you store elements in vectors*.

```{r}
# YOUR CODE HERE
sheep <- as.vector(1:100) #vector 1-100
counting_sheep <- vector (mode = 'integer') #empty integer vector
#multiply each number in sheep by pi and append to counting_sheep
for (i in sheep){
  tmp <- i * pi
  counting_sheep <- append(counting_sheep, tmp)
} 
```
:::

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}
## Task 5

A)  Using the `marine_survey` data frame, use the `lapply` function function to calculate the summary statistics of each numeric column and store this as a `list` called `numeric_summaries`.

B)  Modify the conversion step using `do.call` to flatten this list to a data frame called `numeric_summaries_df`.

C)  Add a column called `variables` using the rownames.

```{r}
#A)
# CODE TO FIX
#numeric_summaries <- vapply(marine_survey(sapply(is.character, marine_survey)), summary)

#create a vector of which column numbers in marine_survey are numeric
numeric_columns <- which(sapply(marine_survey, FUN = is.numeric))

#apply summary function to the columns of marine_survey specified above
numeric_summaries <- lapply(marine_survey[, numeric_columns], FUN = summary)


#B) 
# CODE TO FIX
# numeric_summaries_df <- do.call(cbind, replicate(numeric_summaries, as.data.frame.list))

numeric_summaries_df <- data.frame(do.call(cbind, numeric_summaries)) #that was diabolical

# C) 
#convert rownames to a column called variables
numeric_summaries_df <- rownames_to_column(numeric_summaries_df, 'variables') 

numeric_summaries_df #print final result
```
:::

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}
## Task 6

Rewrite Task 2 in the lecture exercises (summing total abundance for each species, for each year) using group_by() and summarize() instead of a for() loop. Save in an object called `total_abundance`.

```{r}
# Fish abundance data from lecture exercises
fish_abundance <- read.csv("../Data/Groundfish_surveys.csv")
#convert to dataframe
abundance_df <- data.frame(fish_abundance)
```

```{r}
# YOUR CODE HERE
#Group abundance_df by species and save to a new variable
#This doesn't change its appearance but allows summarize() in the next line to work
by_species <- abundance_df %>%  group_by(accepted_name)

#use summarize() to sum all cpue numbers per species (this works in R but doesn't
#seem super biologically sound - can you really sum CPUE?)
by_species %>% summarize(
  total_abundance = sum(num_cpue)
)
```
:::

**BEFORE SUBMITTING**

Clear your environment (broom) and run your entire script to ensure everything has been defined properly!

Be sure to submit this `markdown` file to the relevant Brightspace submission box.
