---
title: "Biol 3872 Lab4-Assignment"
author: "Jesse"
student_id: "420"
date: "`r Sys.Date()`"
editor_options:
  chunk_output_type: inline
---

\
**BEFORE STARTING:**\
1. Include your name (as shown in Brightspace) in the `author` section above as text (e.g., "Liam MacNeil") and Banner ID ("B00123456") in `student_id`.\
2. Install and load the packages required to run your code chunks/

Assignments will largely be evaluated based on your code chunk answers. If they are incomplete or formatted incorrectly (e.g., code chunks don't run), no marks will be awarded. I recommend experimenting with solving these answers in a separate script/markdown document and reporting your answers with minimal script.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load the packages needed to answer the questions below. If you realize you need a package, load (install if needed) here.
library(tidyverse)
library(RColorBrewer)
library(colorspace)
library(patchwork)
```

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}

## Task 1 {2 points}

A) Using the iris dataset, create a `ggplot` showing only the mean sepal length for each species using the `stat_summary` function and `geom` crossbar. Store in variable `iris_mean.` 

B) Now add a `geom` jitter layer to your plot showing how the data is distributed around the mean. Store in variable `iris_data.`
```{r}
data("iris")
#A)
# YOUR CODE HERE
iris_mean <- 
ggplot(data = iris, aes(x = Species, y = Sepal.Length))+
  stat_summary(fun = mean, geom="crossbar")
  
#B)
# YOUR CODE HERE
iris_data <- iris_mean + geom_jitter()
```
:::

\

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}

## Task 2 {3 points}

Use the `mtcars` data set to compare fuel efficiency of car models and store as unique `ggplot` objects for each question named `fuel_efficiency_plot_a`, `fuel_efficiency_plot_b`, etc. Think about Tufte's rules 5, 7, 9.

A)  Transform rownames to a column titled `cartype`, convert this column to a factor ordered (descending) by the average `mpg` (miles per gallon), pipe this modified dataframe into a `ggplot` barplot.

B)  Add to `fuel_efficiency_plot_a`, a minimal theme with 12 font size, adjust the x labels to 45 degrees and horizontally (right) justify labels.

C)  Add to `fuel_efficiency_plot_b`, another column `geom` which is filled by car weight with a `viridis` color scheme.
```{r}
data("mtcars")
# A)
# YOUR CODE HERE
cars.mut <- #write all the following to a new variable
mtcars %>% 
    rownames_to_column(var = "cartype") %>% #make the rownames into a column so we can call it by name
    mutate(cartype = factor(cartype))  #make the new column int a factor

cars.mut$cartype <- fct_reorder(.f = cars.mut$cartype,
                                .x = cars.mut$mpg,
                                .desc = TRUE) #reorder the cartype factor by referencing the mpg column, in descending order

#plot it
fuel_efficiency_plot_a <-
ggplot(cars.mut, aes(x = cartype, y = mpg))+
  geom_col()

fuel_efficiency_plot_a

# B)
# YOUR CODE HERE
fuel_efficiency_plot_b <-
  fuel_efficiency_plot_a +
    theme_minimal(base_size = 12)+
    theme(axis.text.x = element_text(angle=45, hjust=1))
  
fuel_efficiency_plot_b  
  
# C)
# YOUR CODE HERE
fuel_efficiency_plot_c <-
  fuel_efficiency_plot_b +
    geom_col(aes(fill = wt))+
    scale_fill_continuous(palette = viridis(length(cars.mut$wt)))

fuel_efficiency_plot_c
```
:::

\

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}

## Task 3 {1 point}

A)  Use `patchwork` to combine a new plot with `fuel_efficiency_plot_c` called `horsepower_plot`. Execute similar data manipulation steps as in **Task 2** to convert rownames to an ordered column as a factor. Pipe this step into a ggplot object with horsepower as the y axis (Think: What would the x axis need to be to combine with `fuel_efficiency_plot_c`?), add a scatterplot `geom` with enlarged points (`size = 3`), a minimal theme with 12 point font, a modified x axis and color palettes identical to `fuel_efficiency_plot_c`.

Patch this `cars_compared` plot with `fuel_efficiency_plot_c` as a single column with two rows and combine the two x axes to avoid redundancy.

```{r}
# A)
# YOUR CODE HERE

horsepower_plot <-
  ggplot(data = cars.mut, aes(x = cartype, y = hp))+
  geom_point(size = 3, shape = 21, aes(fill = wt))+ #only shapes 21-25 allow fill
  #viridis palette, scaled to car weight
  scale_fill_continuous(palette = viridis(length(cars.mut$wt)))+
  #theme
  theme_minimal(base_size = 12)+
  #angle & justify x-axis labels so they fit
  theme(axis.text.x = element_text(angle=45, hjust=1))

horsepower_plot

#combine the two plots using patchwork
cars_compared <- 
  fuel_efficiency_plot_c +
  horsepower_plot +
  #collect x axis label + scale and color legend since they're duplicates
  plot_layout(ncol = 1, axis_titles = "collect_x", axes = 'collect_x',
              guides = 'collect')

cars_compared
```
:::

\

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}

## Task 4 {2 points}

A) Load in the CO2 data set used during the lab exercise and render it as a data frame called `co2_df` with columns `co2`, `year`, and `month`. Recall the steps taken during the exercise to properly format the data frame. Create a new column called `season` by three month intervals (Dec, Jan, Feb = Winter / Mar, Apr, May = Spring, etc.). Check out the `ifelse` statements from lab 2.

B) Order the season variable as a factor (Levels = Winter, Spring, Summer, Autumn). Create a `ggplot` object named `seasonal_co2` plotting years on the x-axis and co2 concentration on the y-axis. Add a point `geom` colored by season with these colors ("#0072B2", "#009E73", "#CC79A7", "#E69F00"), a jitter `geom` to separate points for visualization, and a minimal `theme` with 12 point font.

```{r}
# A)
data("co2")
# YOUR CODE HERE
#create dataframe with columns for c02 level, yearn and month
co2_df <- data.frame(
  co2 = as.numeric(co2),
  year = trunc(time(co2)),
  month = month.abb[cycle(co2)]
)

#for/if/else statement which adds a season column to the dataframe, and fills it by
#determining season based on month
for (i in 1:nrow(co2_df)) {
  if (co2_df$month[i] %in% c('Dec', 'Jan', 'Feb')){
    co2_df$season[i] <- 'Winter'
    }
  else if (co2_df$month[i] %in% c('Mar', 'Apr', 'May')){
    co2_df$season[i] <- 'Spring'
    }
  else if (co2_df$month[i] %in% c('Jun', 'Jul', 'Aug')){
    co2_df$season[i] <- 'Summer'
    }
#Leaving this silly phrasing for future reference if I need to do something silly
  else if (co2_df$month[i] == 'Sep' | co2_df$month[i] == 'Oct' | co2_df$month[i] == 'Nov') {
    co2_df$season[i] <- 'Autumn'
    }
  else {
    print('panic time something is wrong')
    }
}
  

# B)
# YOUR CODE HERE
#make season an ordered factor
co2_df$season <- factor(co2_df$season, levels = c('Winter', 'Spring', 'Summer', 'Autumn', ordered = TRUE))

#plot co2 by year, colored by season
ggplot(co2_df, aes(x= year, y = co2, col = season))+
  geom_point()+
  geom_jitter()+
  scale_colour_manual(values = c("#0072B2", "#009E73", "#CC79A7", "#E69F00"))

```
:::

\

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}

## Task 5 {1 point}

A) Using the same `co2_df` as above, order the months by the calender order (not alphabetical) as shown in the exercise. Create a `ggplot` object named `ordered_co2_friendly` comparing yearly co2 concentrations (same as Task 4). Add a point `geom` and a minimal `theme` with 12 point font and a manual color-blind friendly palette `scale` using `diverge_hcl` from `colorspace`. Finally, `facet` this plot by season.



```{r}
# A)
# YOUR CODE HERE
ordered_co2_friendly <-
co2_df %>% 
  mutate(month = factor(month, levels = month.abb)) %>%
  ggplot(aes(x = year, y = co2, col = month))+
  geom_point()+
  geom_jitter()+
  theme_minimal(base_size = 12)+
  scale_colour_discrete(palette = diverging_hcl(12, palette = 'Blue-Red'))+
  facet_wrap(vars(season))
  
  
ordered_co2_friendly
```
:::

\

**BEFORE SUBMITTING**

Clear your environment (broom) and run your entire script to ensure everything has been defined properly!

Be sure to submit this `markdown` file to the relevant Brightspace submission box.
