---
title: "Biol 3872 Lab6-Assignment"
output: html_document
date: "`r Sys.Date()`"
author: "jesse"
student_id: "B0042069"
---

\
**BEFORE STARTING:**\
1. Include your name (as shown in Brightspace) in the `author` section above as text (e.g., "Liam MacNeil") and Banner ID ("B00123456") in `student_id`.\
2. Install and load the packages required to run your code chunks/

Assignments will largely be evaluated based on your code chunk answers. If they are incomplete or formatted incorrectly (e.g., code chunks don't run), no marks will be awarded. I recommend experimenting with solving these answers in a separate script/markdown document and reporting your answers with minimal script.

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}
## Task 1 {2 points}

A)  Convert and if needed, correct the character strings to dates 1-5 in unambiguous formats. Store these as variables with the given names.

B)  Convert and if needed, correct the character strings to datetimes 1-3 in unambiguous formats. Store these as variables with the given names. Give all datetimes in Atlantic Standard Time (AST). Check `?OlsonNames`.

```{r}
# YOUR CODE HERE

# A)
# Roald Amundsen reaches the South Pole
date1 <- mdy("Dec/14/1911", tz = 'Canada/Atlantic')

# Alexander Graham Bell and Thomas Watson exchange the first telephone call
date2 <- mdy("January 25, 1915", tz = 'Canada/Atlantic')

# Sputnik is launched into orbit
date3 <- mdy("Oct 4 1957", tz = 'Canada/Atlantic')

# The first Woodstock festival begins
date4 <- mdy("August 16 69", tz = 'Canada/Atlantic')

# The first Blackberry mobile device is released by Research In Motion
date5 <- dym("19/99/January", tz = 'Canada/Atlantic')


# B)


# The estimated minute (Pacific Standard Time) Sidney Crosby scored the overtime goal to win gold in 2010 Olympics
datetime1 <- as.POSIXct("February 28th, 2010 2:38PM", format = '%B %dth, %Y %I:%M%p', tz = 'Canada/Atlantic')


# The timestamp of the first observations detecting gravitational waves. These were produced by the whirling merger of two black holes, generating ripples in space-time confirming Einstein's prediction from general relativity (reported in UTC; observed first at 05:50:45 in Louisiana within Eastern Standard Time)
datetime2 <- as.POSIXct("Sept/14/2015 09:50:45", format = '%bt/%d/%Y %H:%M:%S')

# The beginning minute a total eclipse* was visible over large sections of Nova Scotia
# * These occur every ~18 months, but in any one (inhabited) place, may not occur for centuries.
datetime3 <- as.POSIXct("04-8-2024 16:39", format = '%m-%d-%Y %H:%M')


```
:::

\

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}
## Task 2 {3 points}

A)  Using the Bedford Basin data (`Bedford_ts`) and your skills indexing data, extract all the rows where `year` equals 2016, store this data in a variable called `ts_2016`. **Do you need to load certain packages to complete this?**

B)  Convert the `month` column to full month names in column called `month_full`.

C)  Summarize microbial richness (median) by each year and each month, then plot it using `ggplot` with the x-axis showing full month names (**arranged by calender order**), coloring points by `Year`. Store this plot in a plot object called `monthly_richness`.

**NOT MARKED**: Make this plot pretty: e.g., modify x-axis labels to be readable, edit `Year` to be discrete, apply a nice color palette, and customize in other ways that improve plot quality.

```{r}
library(tidyverse)

Bedford_ts <- read.csv("../Data/BedfordBasin_ts.csv") %>%
  mutate(Year = year(Date),
         Month = month(Date))

# YOUR CODE HERE
# A)
ts_2016 <- subset(Bedford_ts, Year == 2016)

# B)
ts_2016$month_full <- month.name[ts_2016$Month]


# C)

richness <- 
  Bedford_ts %>% 
    group_by(month, year) %>% 
    summarise(Richness = median(Richness))

#make year into a discrete factor
richness$year <- as.factor(richness$year) 
#make a column for full month names
richness$month_name <- as.factor(month.name[richness$month]) 
#reorder month name according to numerical month
richness$month_name <- fct_reorder(.f = richness$month_name,
                                   .x = richness$month)
#plot it
ggplot(data = richness, aes(x = month_name, y = Richness, col = year))+
  geom_point()+
  theme(axis.text.x = element_text(angle=45, vjust=1,
                                   hjust=1))


  


```
:::



::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}

## Task 3 {3 points}

A)  Take the sequences of `day`, `month`, `year`, `hour`, `min`, and `secs` and combine them into a single vector of unambiguous datetimes named `combined_datetime`.

B)  Search `OlsonNames` to find the time zone for Stockholm (Sweden) and assign this to datetimes in a new object called `combined_datetime_stock`.

C)  Convert these times to the Atlantic Canada timezone (AST) and store named these datetimes in `combined_datetime_NS`.
day <- seq(1,30, 1)
month <- rep(seq(1, 10, 1), 3)
year <- 1978:2007
hour <- rep(11:20, 3)
min <- 15:44
sec <- 4:33

# A)


# B)

# C)

```
:::

\

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}
## Task 4 {3 points}

A)  Convert vector of `dates` to a date object called `dates_format`.

B)  Calculate the difference between each date in sequence (date 1 to date 2, date 2 to date 3, etc.) and store this as a **vector** called `dfftime_v1`. This can be done manually, you can also check functions from `lubridate` (`?diff`), but remember to convert to a vector.

C)  Index `dates_format` to remove all dates which are after the date when gravitational waves were first measured (see Q1). Store this as a vector called `pre_Einstein`.

```{r}
dates <- c("1974-01-01", "1981-09-04", "1993-05-18", "2004-12-11", "2015-09-14", "2023-01-13")

# YOUR CODE HERE
# A)

# B)

# C)

```
:::

\

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}
## Task 5 {2 points}

A)  Using the Bedford Basin data (`Bedford_ts`), create a `for()` loop or use an `*apply` that calculates the weekly median microbial richness and store it in a variable called `weekly_median`.

B)  Create a `ggplot` object called `Richness_weekly` which plots raw richness (y-axis) against weeks (x-axis). Color points based on their `month_full` name (ordered by calender). Overlay `weekly_median` into this plot with a unique `geom_point` layer.\

```{r}
# YOUR CODE HERE
# A)


# B)



```
:::

Clear your environment (broom) and run your entire script to ensure everything has been defined properly!

Be sure to submit this `markdown` file to the relevant Brightspace submission box.
